### 2024-10-23 paper-en.tex
Last chapter and theorem reproof

\begin{definition}
    Let $G$ be a metagraph with variable weights. The girth of $G$ is the maximum among all girths of r-extensions of $G$.
\end{definition}

@@

\begin{notice}
    Any complete metagraph that has at least 2 vertices in one part and 3 in another has girth less than or equal to 12. Let us consider metagraph $G$ in figure \ref{full_graph_2_by_3}. $\mu =(e_1, e_4, e_5, e_2, e_3, e_6, e_4, e_1, e_2, e_5, e_6, e_3)$ is a cycle. Its characteristic is euqal to $x_1 - x_4 + x_5 - x_2 + x_3 - x_6 + x_4 - x_1 + x_2 - x_5 + x_6 - e_3 = 0$. Thus paths that correspond to $\mu$ are cycles on r-expansions of $G$.

    Thereby to build r-extensions with girths greater than 12 non complete metagraphs must be used.
\end{notice}

@@

>>>>

%% Проверить, что оно нигде дальше не используется
% \begin{definition}
%     Let $G$ be a metagraph with variable weights. The girth of $G$ is the maximum among all girths of r-extensions of $G$.
% \end{definition}

@@

\begin{notice}
    For any metagraph that has $K_{2,3}$ as a subgraph, the girth of any its r-expansion is at most 12.
    Consider metagraph $G$ shown in figure \ref{full_graph_2_by_3}. Let $\mu = (e_1, e_4, e_5, e_2, e_3, e_6, e_4, e_1, e_2, e_5, e_6, e_3)$ be a cycle. Its characteristic is euqal to $x_1 - x_4 + x_5 - x_2 + x_3 - x_6 + x_4 - x_1 + x_2 - x_5 + x_6 - x_3 = 0$. Therefore, paths that correspond to $\mu$ are cycles with length of $12$ on r-expansions of $G$.
\end{notice}

@@

    Let $G(V, E, f, w)$ be a metagraph, let $\mu$ be a cycle on metagraph $G$, let $a$ and $b$ be two (possibly equal) vertices on cycle $\mu$, and let $\eta$ be a path from $a$ to $b$ such that the first and the last edges of $\eta$ do not belong to $\mu$. Then girth of any r-expansion of $G$ is not greater than $2(|\mu| + |\eta|)$.

\begin{proof}

    Denote inverse paths of $\mu$ and $\eta$ as $\mu'$ and $\eta'$ respectively.

    Denote gluing $\nu = (\eta, \mu, \eta', \mu')$. This edge sequence does not have two equal edges in a row since the first and the last edges of the path $\eta$ do not belong to $\mu$, therefore, $\nu$ is a cycle on $G$.

    Length of $\nu$ is $2(|\mu| + |\eta|)$.

    Applying theorem \ref{glued_paths} calculate characteristic of $\nu$: $\chi(\nu) = \chi(\eta) + \chi(\mu) + \chi(\eta') + \chi(\mu') = \chi(\eta) + \chi(\mu) - \chi(\eta) - \chi(\mu) = 0$.

    Thus corresponding paths of $\nu$ are cycles on any r-expansion of $G$. This means that girth of any r-expansion of $G$ is at most $2(|\mu| + |\eta|)$.

\end{proof}

% Становится полноценной теоремой, которую надо доказать

\begin{theorem}\label{theorem_cycle_expansion}

    Let $G(V, E, f)$ be a bipartite graph with girth $2l$. Let $G'(V, E, f, w')$ be a metagraph that is a weighted graph $G$. Then there is an r-expansion of $G'$ with girth greater then or equal to $6l$.

Thus, to construct graphs with greater girths metagraphs with greater girths should be used. For example multiple arcs must be avoided because they cause cycles with length $2$ significantly limiting the girth of an r-extension.

\textbf{Algorithm 3.}

To build a bipartite graph with girth greater then or equal to given number the following steps can be used:

\begin{itemize}
    \item Build $G_0$ a complete graph, it has a girth of $4$.
    \item Build $G_1$ applying algorithm 2 to metagraph with variable weights with structure from graph $G_0$. The girth of $G_1$ is greater then or equal to $6$.
    \item Repeat the previous step applying algorithm 2 to graph with greater and greater girths, until the graph with the required girth is constructed.  
\end{itemize}

Use algorithm 3 to build a (2,3)-regular bipartite graph with girth $14$.

Build a (2,3)-regular complete metagraph. It has 2 control and 3 informational vertices and has girth of $4$. This graph is shown in figure \ref{full_graph_2_by_3}.

Bipartite graph with girth $8$ build with algorithm 2 is shown in figure \ref{be_not_afraid}. It has 6 control and 9 informational vertices. Control vertices are located in the inner circle and informational on the outer one. Vertices are grouped by their corresponding vertex in metagraph.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.2]{graph_2.png}
    \caption{ (2,3)-regular bipartite graph with girth of 8. }
    \label{be_not_afraid}
\end{figure}

Bipartite graph with girth $14$ is shown in figure \ref{be_not_afraid_2}. It has 42 control and 63 informational vertices.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.2]{graph_3.png}
    \caption{ (2,3)-regular bipartite graph with girth of 14. }
    \label{be_not_afraid_2}
\end{figure}

### 2024-10-23 paper.tex
Last chapter and theorem reproof

\textbf{Следствие.}

Если обхват двудольного графа ~-- $2l$, то обхват метаграфа с неопределёнными весами, построенного на основе этого графа будет не меньше $6l$.

Таким образом, для построения графов с большим обхватом следует выбирать метаграфы с большим обхватом. Например следует избегать кратных дуг в метаграфе, так как кратная дуга порождает цикл длины 2 и может значительно снизить обхват r-расширения.

>>>>

\textbf{Теорема.}

Пусть $G'$ ~--- двудольный граф. Пусть $\Phi$ ~--- множество всех структур вида "цикл с ручкой" на графе $G'$. Обозначим $\theta = \min\limits_{\phi \in \Phi} \{ |\phi| \}$. Тогда существует набор весов дуг графа $G'$ таким образом, чтобы для полученного метаграфа существовало расширение с обхватом больше или равным $2 \theta$.

Предположим, что на всех расширениях для любых наборов весов существуют циклы длины меньшей $2\theta$.

Рассмотрим такие циклы на метаграфе, которым соответствуют циклы длины меньшей $2 \theta$ только на некоторых расширениях.

Так как существуют расширения на которых таким циклам соответствую пути не являющиеся циклами, можно используя Алгоритм 1 подобрать веса и $r$ такие что всем таким циклы будут соответствовать не циклы на расширениях. Таким образом необходимо рассмотреть только циклы, которым соответствуют циклы на всех расширениях.

Если циклу на метаграфе соответствуют циклы на всех расширениях, значит его характеристика обращается в 0 вне зависимости от выбранных весов. Значит такой цикл проходит каждую свою дугу равное количество раз в противоположные стороны.

Путь не может проходить одну и ту же дугу дважды подряд, значит его структура распадается на два цикла, которые проходятся в противоположные стороны. Эта структура соответствует циклу с ручкой, следовательно на метаграфе существует цикл с ручкой длины меньшей или равной $(2\theta - 2) / 2 = \theta - 1$, что противоречит условию теоремы. Следовательно существует расширение, на котором нет цикла длины меньшей или равной $2\theta - 2$.
